---
title: 'R Programming | Course #2 of Data Science Specialization by Johns Hopkins'
author: "Hira Tariq"
date: "2023-06-24"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Week 1: R Data_Types

* **Vectors** Declaration: `a <- c(1,2,3,5)`; Indexing: `a[n]`

* **Matrices** `m <- matrix(1:4, nrow = 2, ncol = 2)`; Indexing: `matrix[row, col]`
<input type=button class=hideshow></input>
  + Matrices are vectors with Dimension attributes (so we can use dim() function on matrices as well) 
  + Matrix Declaration (Note: Matrix is constructed Column-wise, columns are filled first)
  + Matrix Attributes & Indexing
  + We can create Matrix from a vector as well by creating dimensions after declaration
  + We can create Matrix through Column and Row Binding of vectors using cbind(); rbind functions


```{r,  collapse = TRUE}
m <- matrix(1:4, nrow = 2, ncol = 2)
m
attributes(m)
# Dim() outputs a vector of length 2 
m[1,1]
m[1,]

Vec <- 1:10
# now adding dimensions
Vec
dim(Vec) <- c(2,5)
Vec

c1 <- 1:3
c2 <- 13:15
matrix_c12 <- cbind(c1,c2)
matrix_c12
attributes(matrix_c12)
matrix_r12 <- rbind(c1,c2)
matrix_r12
attributes(matrix_r12)
```

* **Factors** to store Categorical Data; has labels underneath; has use cases in modeling funcions like lm(), glm() etc.

```{r, collapse = TRUE}
x <- factor(c("Male", "Female", "Male", "Female", "Female", "Female"))
print(x)
table(x)

unclass(x)

x <- factor(c("Male", "Female", "Male", "Female", "Female", "Female"), levels = c("Male", "Female"))
x
table(x)
unclass(x)

```
* Remove NA/ Missing Values --> **is.na()** & sub-setting through **complete.cases()** function
<input type=button class=hideshow></input>
  + **is.na(x)**: outputs logical vector of TRUE FALASE; TRUE wherever there is a missing value --> !is.na() will convert FALSE to TRUE
```{r, collapse= TRUE}
x <- c(1,2,4, NA, NaN, 5, NA, NULL, 6)
is.na(x)
# Notice: Here NULL is not being treated as a missing value or anything, it's just not part of the vector thus vector size here is 8 instead of 9
length(x)
# Use cases of is.na() to exclude missing values 
x[is.na(x)]
x[!is.na(x)]

# Missing Value Count
length(x[is.na(x)])
# Length of x excluding missing values
length(x[!is.na(x)])
```
  + **complete.cases(x,y)**: will output TRUE where x & y both have non-missing values -> use cases could be where we are trying multiple features in the model and we want to check for their correlations may be
```{r, collapse=TRUE}
# --------------------------- Example of Subsetting out Missing Values from the Dataset ------------------------------
x <- c(1,2,3,NA, NaN, 6, 7)
y <- c(1, NA, 3, 4, NaN, 6, NA)
complete.cases(x,y)
x[complete.cases(x,y)]
y[complete.cases(x,y)]

airquality[1:6,]
complete.cases(airquality[1:6,])
airquality[complete.cases(airquality[1:6,]),]
airquality[complete.cases(airquality[1:6,]),][1:6,]

```
  




<script>
$( "input.hideshow" ).each( function ( index, button ) {
  button.value = 'Hide';
  $( button ).click( function () {
    var target = this.nextSibling ? this : this.parentNode;
    target = target.nextSibling.nextSibling;
    if ( target.style.display == 'block' || target.style.display == '' ) {
      target.style.display = 'none';
      this.value = 'Show';
    } else {
      target.style.display = 'block';
      this.value = 'Hide';
    }
  } );
} );
</script>
